---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wellr

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/wellr)](https://CRAN.R-project.org/package=wellr) [![R-CMD-check](https://github.com/bradyajohnston/wellr/workflows/R-CMD-check/badge.svg)](https://github.com/bradyajohnston/wellr/actions) [![Codecov test coverage](https://codecov.io/gh/bradyajohnston/wellr/branch/master/graph/badge.svg)](https://app.codecov.io/gh/bradyajohnston/wellr?branch=master)

<!-- badges: end -->

`{wellr}` provides a consistent and reliable interface for dealing with plate-based data and related calculations.
It provides functions for reading the output of various plate readers, indexing through and reformatting microtitre plates, converting between well IDs ("C05") and their respective rows, columns and index.

## Installation

<!-- You can install the released version of wellr from [CRAN](https://CRAN.R-project.org) with: -->

### From R-Universe:
``` r
install.packages("wellr", repos = "bradyajohnston.r-universe.dev")
```

### From GitHub
``` r
# install.packages("remotes")
remotes::install_github("bradyajohnston/wellr")
```


## Basic Examples

```{r example}
library(wellr)

well_format("G8")
well_to_col_num("G8")
well_to_row_num("G8")
well_to_index("H1")
well_to_index("H1", colwise = TRUE)
well_from_index(37)
well_from_index(37, colwise = TRUE)
well_join(3, 8)
well_join("E", 10)
```

## Reading Biotek

Get the file paths of the demo files.

```{r}
file_data <- system.file("extdata",
  "20220929_1steptimer20.csv",
  package = "wellr"
)

file_meta <- system.file("extdata",
  "20220929_1steptimer20_metainfo.csv",
  package = "wellr"
)
```

Read in an example plate from a Biotek plate reader.

```{r}
plate <- plate_read_biotek(file_data)
plate
```

```{r}
plate |>
  plate_add_meta(file_meta)
```

## Creating Dummy Plates

Create a data frame for plate-based data.

```{r}
well_plate(8, 12)
```

## Helpful Plotting Functions

```{r}
set.seed(3)
plate <- well_plate(8, 12)[, "well"]
plate$value <- rnorm(96, sd = 10)

well_plot(plate, well, value)
```
