---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# wellr

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/wellr)](https://CRAN.R-project.org/package=wellr) [![R-CMD-check](https://github.com/bradyajohnston/wellr/workflows/R-CMD-check/badge.svg)](https://github.com/bradyajohnston/wellr/actions) [![Codecov test coverage](https://codecov.io/gh/bradyajohnston/wellr/branch/master/graph/badge.svg)](https://app.codecov.io/gh/bradyajohnston/wellr?branch=master)
[![R-CMD-check](https://github.com/BradyAJohnston/wellr/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/BradyAJohnston/wellr/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

A set of _tidy_ utilities for working with plate based data. Wet-lab experiments are often carried out in microplates of varying sizes. `{wellr}` aims to provide a cleaner interface to working with data from plate readers that works well within the `{tidyverse}` [design principles](https://design.tidyverse.org/unifying.html).


## Installation

<!-- You can install the released version of wellr from [CRAN](https://CRAN.R-project.org) with: -->

### From R-Universe:
This is the recommended way to install the package - as it requires no additional dependencies. It is currently not available on CRAN.
``` r
install.packages("wellr", repos = "bradyajohnston.r-universe.dev")
```

### From GitHub
``` r
# install.packages("remotes")
remotes::install_github("bradyajohnston/wellr")
```


## Basic Examples

```{r example}
library(wellr)

well_format("G8")
well_to_col_num("G8")
well_to_row_num("G8")
well_to_index("H1")
well_to_index("H1", colwise = TRUE)
well_from_index(37)
well_from_index(37, colwise = TRUE)
well_join(3, 8)
well_join("E", 10)
```

## Reading Biotek

Get the file paths of the demo files.

```{r}
file_data <- system.file("extdata",
  "20220929_1steptimer20.csv",
  package = "wellr"
)

file_meta <- system.file("extdata",
  "20220929_1steptimer20_metainfo.csv",
  package = "wellr"
)
```

Read in an example plate from a Biotek plate reader.

```{r}
plate <- plate_read_biotek(file_data)
plate
```

```{r}
plate |>
  plate_add_meta(file_meta)
```

### Read Biotek Wavelength Data

If the biotek `.csv` file includes spectral readings from different wavelengths, these won't be included in the regular `plate_read_biotek()` function's output - as they don't have associated time information. 

The `plate_read_biotek_wl()` function extracts these readings, and the resulting data frame includes a `id` column, specifying which wavelength reading they come from.

```{r}
file_including_wavelength <- system.file(
  "extdata", "2024-02-29_vio_GFP_main.csv",
  package = "wellr"
)

plate_read_biotek(file_including_wavelength)

```
If reading in fluorescent data, there will sometimes be two wavelengths reported in the `.csv`. By default just the first wavelength will be used for the column names, but you can ensure that both wavelengths are included by useing `second_wl = TRUE`. The `gfp` column now includes the second wavelength that the fluorescence was measured at.
```{r}
plate_read_biotek(file_including_wavelength, second_wl = TRUE)
```

When reading the spectral data from the file, the readings do not include time data.

The resulting data frame will however include a `id` column with `id = 1` being the first reading, `id = 2` being the second reading etc.
```{r}
plate_read_biotek_wl(file_including_wavelength)
```


## Creating Dummy Plates

Create a data frame for plate-based data.

```{r}
well_plate(8, 12)
```

## Helpful Plotting Functions

```{r}
set.seed(3)
plate <- well_plate(8, 12)[, "well"]
plate$value <- rnorm(96, sd = 10)

well_plot(plate, well, value)
```
